---
title: "gWidgets2, a GUI system for R"
author: David C Sterratt
date: 20th January, 2016
output: ioslides_presentation
---

## What is gWidgets2?

- gWidgets2 by John Verzani is a reworking of the earlier gWidgets package
- Provides a toolkit-independent API for building interactive GUIs on Windows, Mac and Linux
- GTK2/3, Tk, Qt
- At present GTK is the best supported toolkit on the three operating systems
- Requires gWidgets2RGtk2 and RGtk2 (Michael Lawrence)
- Programatic design
  - No GUI designer a la Matlab
- API of gWidgets and gWidgets2 almost identical

# Installation

## Linux

- Prerequisites: GTK2 development libraries
- On Debian/Ubuntu at the command line:
```
sudo apt-get install libgtk2.0-dev
```
- In R:
```{r, eval=FALSE}
install.packages(c("gWidgets2RGtk2", "cairoDevice"))
## May need dev version
devtools::install_github("jverzani/gWidgets2RGtk2")
```

## Windows

```{r, eval=FALSE}
install.packages(c("gWidgets2RGtk2", "cairoDevice"))
library(gWidgets2RGtk2)
```
- There will be an error message saying that libatk-1.0-0.dll ismissing from your computer.
- Click  on OK
- Another  message  will  appear,  asking  if  you  want  to  install GTK+.
- Agree to this.
- After this has finished, quit R (without saving the workspace image) and restart.
    
## Mac

- Pain! - GTK libraries not easy to install
- Options:
  - GTK binaries (should work for pre-Mavericks MacOS X)
  - MacPorts
  - HomeBrew
- More details at: https://github.com/davidcsterratt/retistruct/issues/4

## Concepts

- __Widgets__: buttons, radio buttons, dropdown lists,  sliders, edit boxes ...
- __Container__: where to place a widget
- __Handler__: how a widget responds to events, e.g. button presses
- __Method__: How to get and set widget values

# Hello World

## Hello World

```{r, eval=FALSE}
library(gWidgets2)
## Select the RGtk2 toolkit
options(guiToolkit="RGtk2")

## Create a window called "Hello World"
win <- gwindow("Hello World")

## Create a "Hello world" button and make it do something using a
## handler
obj <- gbutton("Hello world", container=win,
               handler=function(h, ...) {
                 gmessage("Hello world")
               })
```

- Note *widgets* (gbutton, gmessage), *container* and *handler*
- Handler function has to be of form `function(h, ...)`

## Goodbye world - groups

```{r, eval=FALSE}
library(gWidgets2)
options(guiToolkit="RGtk2")
win <- gwindow("Hello World, Goodbye world")

## Create a ggroup to  put more than one Widget in
group <- ggroup(horizontal=FALSE, container=win)

## Create a "Hello world" button within the group we've just made
hello <- gbutton("Hello world", container=group,
               handler=function(h, ...) {
                 gmessage("Hello world")
               })
## Add another button
goodbye <- gbutton("Goodbye world", container=group,
                   handler=function(h, ...) {
                     gmessage("Goodbye world")
               })
```

# Methods

## Hello World x n

```{r, eval=FALSE}
library(gWidgets2)
options(guiToolkit="RGtk2")
win <- gwindow("Hello World times n")
group <- ggroup(horizontal=FALSE, container=win)

## Add a spinner in a group
g.times <- ggroup(horizontal=TRUE, container=group)
text <- glabel("How many times?", container=g.times)
times <- gspinbutton(from=1, to=5, by=1, value=3, container=g.times)

## The handler function calls the svalue method of the gspinbutton to
## get the number of times
hello <- gbutton("Hello world", container=group,
                 handler=function(h, ...) {
                   gmessage(rep("Hello world", svalue(times)))
                 })
```

## Including graphics
```{r, eval=FALSE}
library(gWidgets2)
options(guiToolkit="RGtk2")
## cairoDevice allows plotting in GTK ggraphics widgets
library(cairoDevice) 
win <- gwindow("Hello World, the next generation")
group <- ggroup(horizontal=FALSE, container=win)

## Create an area for a graphics device
g.d <- ggraphics(container=group, width=600, height=600)
## Create the device in it
d <- dev.cur()
# Set the handler to plot
obj <- gbutton("Plot", container = group,
               handler=function(h, ...) {
                 x <- rnorm(50)
                 plot(density(x, adjust=1, kernel="gaussian"),
                      main="Density plot")
                 rug(x)
               })
```

## Graphics with a method

```{r, eval=FALSE}
updatePlot <- function(h,...) {
  x <- rnorm(50)
  plot(density(x, adjust = svalue(bandwidthAdjust),
               kernel = "gaussian"),main="Density plot")
  rug(x)
}
## Add a slider to store a value
bandwidthAdjust <- gslider(from=0,to=2,by=.01, value=1, 
                           handler=updatePlot, container=group)

```

# Wrapping code

## Hide variables in function

```{r, eval=FALSE}
library(gWidgets2)
## Select the RGtk2 toolkit
options(guiToolkit="RGtk2")
myapp <- function() {
  ## This function is essentially the constructor for a class. The
  ## environment 'this' contains all member variables and functions of
  ## the class. We do not need to refer to 'this' in the code, but we
  ## will return it to facilitate debugging. 
  this <- environment()
  win <- gwindow("Hello World")
  obj <- gbutton("Hello world", container = win,
               handler=function(h, ...) {
                 gmessage("Goodbye world")
               })
  return(invisible(list(env=this)))
}
```

# Demo - Retistruct

# Further Reading

gWidgets vingette

<!--  LocalWords:  gWidgets Sterratt th ioslides Verzani API GUIs GTK
 -->
<!--  LocalWords:  Tk RGtk RGtk Matlab sudo libgtk dev eval jverzani
 -->
<!--  LocalWords:  cairoDevice workspace pre MacOS MacPorts HomeBrew
 -->
<!--  LocalWords:  dropdown guiToolkit gwindow gbutton gmessage rnorm
 -->
<!--  LocalWords:  ggroup glabel gspinbutton svalue ggraphics
 -->
<!--  LocalWords:  gaussian Retistruct
 -->
