---
title: "gWidgets2, a GUI system for R"
author: David C Sterratt
date: 20th January, 2016
output: ioslides_presentation
---

## What is gWidgets2?

- gWidgets2 by John Verzani is a reworking of the earlier gWidgets package
- Provides a toolkit-independent API for building interactive GUIs on
  Windows, Mac and Linux
- GTK2/3, Tk, Qt
- At present GTK is the best supported toolkit on the three operating
  systems
- Requires gWidgetsRGtk2 and RGtk2 (Michael Lawrence)
- Programatic design
  - No GUI designer a la Matlab

# Installation

## Linux

- Prerequisites: GTK2 development libraries
- On Debian/Ubuntu at the command line:
```
sudo apt-get install libgtk2.0-dev
```
- In R:
```{r, eval=FALSE}
install.packages(c("gWidgets2RGtk2", "cairoDevice"))
```

## Windows

```{r, eval=FALSE}
install.packages(c("gWidgets2RGtk2", "cairoDevice"))
library(gWidgets2RGtk2)
```
- There will be an error message saying that libatk-1.0-0.dll is missing
  from  your  computer.
- Click  on OK
- Another  message  will  appear,  asking  if  you  want  to  install
  GTK+.
- Agree to this.
- After this has finished, quit R (without saving the workspace image) and
  restart.
    
## Mac

- Pain! - GTK libraries not easy to install
- Options:
  - GTK binaries (should work for pre-Mavericks MacOS X)
  - MacPorts
  - HomeBrew
- More details at:
  https://github.com/davidcsterratt/retistruct/issues/4

## Concepts

- __Widgets__: buttons, radio buttons, dropdown lists,  sliders, edit boxes ...
- __Container__: where to place a widget
- __Handler__: how a widget resopnds to events, e.g. button presses
- __Action__: extra information passed to a handler
- __Method__: How to get and set widget values

# Hello World

## Hello World

```{r, eval=FALSE}
library(gWidgets2)
## Select the RGtk2 toolkit
options(guiToolkit="RGtk2")

## Create a window called "Hello World"
win <- gwindow("Hello World")

## Create a "Hello world" button and make it do something using a
## handler
obj <- gbutton("Hello world", container=win,
               handler=function(h, ...) {
                 gmessage("Hello world")
               })
```

- Note *widgets* (gbutton, gmessage), *container* and *handler*
- Handler function has to be of form `function(h, ...)`

## Goodbye world - groups

```{r, eval=FALSE}
library(gWidgets2)
options(guiToolkit="RGtk2")
win <- gwindow("Hello World, Goodbye world")

## Create a ggroup to  put more than one Widget in
group <- ggroup(horizontal=FALSE, container=win)

## Create a "Hello world" button within the group we've just made
hello <- gbutton("Hello world", container=group,
               handler=function(h, ...) {
                 gmessage("Hello world")
               })
## Add another button
goodbye <- gbutton("Goodbye world", container=group,
                   handler=function(h, ...) {
                     gmessage("Goodbye world")
               })
```

# Methods

## Hello World x n

```{r, eval=FALSE}
library(gWidgets2)
options(guiToolkit="RGtk2")
win <- gwindow("Hello World times n")
group <- ggroup(horizontal=FALSE, container=win)

## Add a spinner in a group
g.times <- ggroup(horizontal=TRUE, container=group)
text <- glabel("How many times?", container=g.times)
times <- gspinbutton(from=1, to=5, by=1, value=3, container=g.times)

## The handler function calls the svalue method of the gspinbutton to
## get the number of times
hello <- gbutton("Hello world", container=group,
                 handler=function(h, ...) {
                   gmessage(rep("Hello world", svalue(times)))
                 })
```

# Including graphics

# Actions

# Wrapping code

# Demo - Retistruct

